---
title: "HELPrct (Data Verbs) Activity"
author: "Mohil Patel"
date: "March 2, 2022"
output: html_notebook
---


## Set Up:

```{r}
# load packages
library(tidyverse)  # for data verbs like `group-by()` & `summarise()`
library(mosaicData) # includes `HELPrct` data set 
library(mosaic)     # includes `mplot()` interactive plot builder
library(esquisse)   # includes `esquisser()` interactive plot builder


# Load the HELPrct data from `mosaicData` into our RStudio environment
data("HELPrct", package = "mosaicData")

```


## Data Description

```{r}
glimpse(HELPrct)

```

*TASK 1: Write a couple of sentences describing the `HELPrct` data--What is the setting for the data? What are they about?* 

The data is looking at individuals who have struggled with drug abuse. A case is a person who struggles with drug abuse. The variables help describe each individuals drug abuse situation as well as their life situation


## Basic Summaries with `summarise()` 

*TASK 2: Produce an expression involving `summarize()` and `HELPrct` that will produce the following.*

1. *number of people (cases) in `HELPrct` study*
2. *combined total number of times in the past 6 months all the people in `HELPrct` entered a detox program (measured at baseline)* 
3. *mean time (in days) to first use of any substance post-detox for all the people in `HELPrct`*


```{r}
HELPrct %>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE) ) 

```
```{r}
HELPrct %>%
  summarise( use_mean = mean(daysanysub,na.rm=TRUE) )


```


## Subgrouping with `group_by()` 

*TASK 3: Repeat task 2 above, but calculate the results group-by-group.  Show your code and write a sentence or two about what you observe in the results for each of the following:* 

1. *males versus females*

```{r}
HELPrct %>%
  group_by(female)%>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE) ) 
```


2. *homeless or not*

```{r}
HELPrct %>%
  group_by(homeless)%>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE) ) 
```

3. *substance*

```{r}
HELPrct %>%
  group_by(substance)%>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE) ) 
```

4. *break down the homeless versus housed further, by sex*

```{r}
HELPrct %>%
  group_by(female,homeless) %>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE))
```

5. *break down the homeless versus housed further, by substance*

```{r}
HELPrct %>%
  group_by(homeless,substance)%>%
  summarise( total_people = n(),total_detox= sum(e2b,na.rm=TRUE),use_mean = mean(daysanysub,na.rm=TRUE) ) 
```


## TASK 4:

*Include one or more interesting plots of the data involving at least 3 variables per plot. Write a few sentences to explain the story that your plot tells about these data.  You can use one of the relationships that you studied in Task 2, or you can explore a different group of variables in the `HELPrct` that show something interesting.*  

The graph below shows the relationship between the age and drug risk for each of the substances. As you can seee that heroin is a very risky drug to take at any stage as compared to alcohol and cocaine
```{r}
ggplot(data = HELPrct, aes(x = age, y = drugrisk)) + geom_point()  + facet_wrap(~substance, ncol = 4) + labs(title = "")

```


