---
title: "Activity: Popular Names"
author: "Mohil  Patel"
date: "2/22/2022"
output: html_notebook
---



# Set Up

```{r}
# load necessary packages
library(tidyverse)
library(mosaic)
library(dcData)
```


```{r}
# Load the BabyNames data into RStudio environment (from dcData package)
data("BabyNames")

```

# Activity

**Goal**: By the end of the activity, each person will create a graph showing ups and downs in the popularity of four names according to the U.S. social security administration  


### Step 1

```{r}
glimpse(BabyNames)

```
```{r}

BabyNames %>%
  filter(name =="Franklin" )


```

### Step 3

The variable sex from the raw data table do not appear at all in the graph.

The variable count might have been transformed into Popularity. Also Name and year remain as the same in the graph.



### Step 4

Yes, the Names have been filtered to provide only Males names:
Abraham
Franklin
Woodrow

And then group by sex. 

The new Variable of Population has been introduced which corresponds with the count variable.



### Step 5
 Filter the Names  and that they are male.
 As well as grouping done by sex to collapse the sex column. 
 


### Step 7: modify the starter code given in the book to produce your own R code..

```{r}
GlyphReadyForm <- BabyNames %>%
  filter(name %in% c("David", "James", "Mohil"), sex == "M") %>%
  group_by(sex)


head(GlyphReadyForm)
```

```{r}
GlyphReadyForm %>% 
  ggplot(aes(x = year, y = count, group = name)) +
  geom_line( size = 1, alpha = 0.5, aes(color = name)) +
  ylab("Popularity") + xlab("Year")


```